\section{Chiral perturbation theory}
\label{section:chiral pertubation theory}
This section is base on \cite{Schwartz:QFT,weinberg_1996_vol2,Scherer2002IntroductionTC}, in addition to the original work~\cite{Gasser-Leutwyler:chiral,WeinbergPhenom,Scherer:PhysRevD.53.315}

In this paper, we will consider the interaction of the two lightest quarks, the up and down quarks $u$ and $d$, i.e. $N_f = 2$.
In this case, the symmetry group of rotations in the flavor indices are $G_f = U(1) \times \SU(2)_L \times \SU(2)_R$.
The generators of $\SU(2)$ are $T_\alpha = \tau_\alpha / 2$, where $\tau_\alpha$ are the Pauli matrices, as described in \autoref{section:algebra bases}.
The mass matrix of the quarks is
\begin{equation}
    \label{Mass matrix}
    m =
    \begin{pmatrix}
        m_u & 0 \\
        0 & m_d
    \end{pmatrix}
    = \frac{1}{2} (m_u + m_d) \one + \frac{1}{2}(m_u - m_d) \tau_3,
\end{equation}
where $m_u \approx 2.16 \, \text{MeV}$, and $m_d \approx 4.67 \, \text{MeV}$~\cite{PDG}.
This means that the $G_f$ symmetry is \emph{explicitly} broken.
% In last section, we saw that for $m_u = m_d$, the subgroup $U(1) \times \SU(2)_V$ remains intact.
% This is called the isospin-limit.
% However, as $m_u \neq m_d$, also the isospin subgroup $\SU(2)_V$ is broken.
% The difference between these masses is small, which is why isospin is a good quantum number.
Even though the underlying symmetry is only approximate, we can still apply the formalism from Goldstone's theorem and the CCWZ construction by including a small mass term for the Goldstone bosons, which in this case are called \emph{Pseudo Goldstone bosons}.

We now focus on the subgroup $G = \SU(2)_L \times \SU(2)_R$.
The ground state spontaneously breaks this approximate symmetry of the two-flavor QCD Lagrangian.
As quarks $q$ are spinors, a non-zero expectation value of the quark field would break Lorentz-invariance.
Instead, the spontaneous symmetry breaking is characterized by the \emph{scalar quark condensate}, $\ex{\bar q q}$.
The scalar quantity $\bar q q$ is invariant under isospin transformations $H = \SU(2)_V$, but not under $\SU(2)_A$.
The Goldstone manifold, therefore, is $G/H = \SU(2)_L\times \SU(2)_R/\SU(2)_V = \SU(2)_A$.
To model the low-energy dynamics of QCD, we start with the massless QCD Lagrangian,
\begin{align*}
    \Ell^0_\mathrm{QCD}[q, \bar q, A_\mu] 
    = i \bar q \slashed{D} q - \frac{1}{4}G_{\mu \nu}^\alpha G^{\mu \nu}_\alpha.
\end{align*}
Following~\cite{Scherer2002IntroductionTC,Gasser-Leutwyler:chiral}, we couple quarks to external currents.
These can be used to model external fields or capture the symmetry-breaking mass term.
As found in \autoref{section:QCD}, the conserved currents are
\begin{equation}
    V_a^\mu = \frac{1}{2} \bar q \gamma^\mu \tau_a q, \quad
    A_a^\mu = \frac{1}{2} \bar q \gamma^\mu \gamma^5 \tau_a q, \quad
    J^\mu = \bar q \gamma^\mu q.
\end{equation}
In addition, external currents can couple to the scalar and pseudo-scalar quark bilinears
\begin{equation}
    \bar q q, \quad \bar q \gamma^5 q, 
    \quad \bar q \tau_a q, \quad \bar q \gamma^5 \tau_a q.
\end{equation}
The Lagrangian of the external is
\begin{align}
    \Ell_\text{ext} 
    & = 
    - (\bar q q )\, s_0 + (i \bar q \gamma^5 q) \, p_0
    - (\bar q \tau_a q)\, s_a + (i \bar q \gamma^5 \tau_a q) \, p_a
    + \frac{1}{3} J_\mu v_{(s)}^\mu 
    + V_\mu^a v_a^\mu + A_\mu^a a_a^\mu \\
    & = 
    - \bar q \left(s + i \gamma^5 p \right) q
    + \bar q \left(\frac{1}{3} v^\mu_{(s)} + v^\mu + a^\mu \gamma^5\right) \gamma_\mu q.
\end{align}
Here, $v, v_{(s)}, a, s$ and $p$ are the external source currents, where we denote
\begin{equation}
    s = s_0 \one + s_a \tau_a, \quad
    p = p_0 \one + p_a \tau_a, \quad
    v^\mu = \frac{1}{2} v_a^\mu \tau_a, \quad
    a^\mu = \frac{1}{2} a_a^\mu \tau_a.
\end{equation}
Setting $v = v_{(s)} = a = s = p = 0$, we recover the chiral limit.
To include the effect of the quark masses, we set $s = m$.
We denote all the external currents as $j = (v, v_{(s)}, a, s, p )$.
As with the conserved currents, we define
\begin{equation}
    v_\mu = \frac{1}{2}(r_\mu + l_\mu),
    \quad
    a_\mu = \frac{1}{2}(r_\mu - l_\mu).
\end{equation}
With this, as well as 
\begin{equation}
    \bar q (s - i \gamma^5 p) q
    = \bar q_R (s - i p) q_L + \bar q_L (s + i p) q_R,
\end{equation}
we can write the Lagrangian as
\begin{equation}
    \Ell_\text{ext} 
    = - \bar q_R (s + i p) q_L - \bar q_L (s - i p) q_R
    + \frac{1}{3} (J_R^\mu + J_V^\mu)(v_{(s)})_\mu
    + R_\mu^a r^\mu_a + L_\mu^a l^\mu_a
\end{equation}
The generating functional is then
\begin{equation}
    Z[j] 
    = 
    \int \D \bar q \D q \D A_\mu \, 
    \exp{
        i \int \dd^4 x 
        \left( 
            \Ell^0_\text{QCD}[q, \bar q, A_\mu] + \Ell_\text{ext}[q, \bar q, j]
        \right)
    }
\end{equation}
We will now use the CCWZ construction and effective field theory to construct the effective Lagrangian, which obeys
\begin{equation}
    Z[j] = \int \D \pi \, \exp{i \int \dd^4 x \, \Ell_\text{eff}[\pi]},
\end{equation}
where $\pi$ are the Goldstone bosons.

\subsection*{Weinberg's power counting scheme}
\label{subsection:power counting}

We now have a relationship between the underlying Lagrangian and the effective Lagrangian, of the form \autoref{integrating out degrees of freedom}.
However, as discussed earlier, we cannot solve QCD at low energies and thus have no way to integrate out the high energy degrees of freedom.
Instead, we have to make use of Weinberg's ``theorem''.
Using the CCWZ construction, we can obtain the most general Lagrangian obeying the symmetries of QCD. 
In this case, we have not laid any constraints on the theory but the most basic assumptions.
This, however, will result in a theory with an infinite number of free parameters, making it rather unwieldy.
To amend this, we need a scheme to order the terms to compute observable perturbatively.
We are working in the low-energy regime, so it is natural to expand in pion momenta.
As we saw in \autoref{seciton:ccwz construction}, the terms in the Lagrangian will be made up of combinations of the terms $e_\mu$ and $d_\mu$ of the Maurer-Cartan form, $i\Sigma \partial_\mu \Sigma = e_\mu + d_\mu$.
Therefore, all terms in the effective Lagrangian will be proportional to a certain number of derivatives, which Lorentz invariance demands to be even.

Consider the matrix element $\mathcal M$ for a given Feynman diagram with external pion lines with momenta $q_n$, where both the energies and momenta are less or equal to some energy scale $Q$.
If we scale $Q\rightarrow tQ$, and consequently also the external momenta $q_n \rightarrow tq_n$, momentum conservation at each vertex ensures that each internal momentum $p$ of the diagram scales as $p \rightarrow tp$.
Assume this diagram is made up of $V_i$ copies of the vertex $i$, which contain $d_i$ derivatives.
Each of these vertices then scale as $t^{d_i}$.
The propagators contribute a factor $p^{-2}$ and will therefore scale as $t^{-2}$, and the integration measure $\dd^4 p$ scales as $t^4$.
This means that a matrix element with $L$ loops and $I$ internal lines scales as
\begin{equation}
    \mathcal M(q) \rightarrow \mathcal M(t q) = t^D \mathcal M(q),
\end{equation}
where 
\begin{equation}
    D = \sum_i V_i d_i - 2 I + 4 L.
\end{equation}

$D$ is called the \emph{chiral dimension} of $\mathcal M$.
Using the formula \cref{Number of loops} for number of loops in a Feynman diagram, we get
\begin{equation}
    D = \sum_i V_i(d_i - 2) + 2 L + 2.
\end{equation}
For low energy scales $Q$, the larges contribution will come from the matrix elements of the smallest chiral dimension $D$.
A general process will consist of a sum of matrix elements of different chiral dimensions.
We can expand this element in powers of the pion momenta by using $t$ as the expansion parameter.
The leading order term will be those where $L = 0$ and $d_i = 2$ so that $D = 2$.
This means that all tree-level contributions to lowest chiral dimension are from terms in the Lagrangian with exactly two derivatives.
Next is $D = 4$, which contains both tree-level contributions from terms with $d_i = 4$ and a one-loop contribution from $d_i = 2$.
We therefore expand the effective Lagrangian as
\begin{equation}
    \Ell_\text{eff} = \Ell_2 + \Ell_4 + ...,
\end{equation}
where $\Ell_{2n}$ contains $2n$ derivatives.
This is equivalent to scaling the space-time cooridnates as $x^\mu \rightarrow tx^\mu$, and expanding the Lagrangian in powers of $t$.

We must also allow for the fact that pions have non-zero mass and that there is a finite isospin chemical potential.
This is all implemented by external currents, which we also have to assume are small.
Any external pions are on-shell, so the pion mass $m_\pi$ must be less than the energy scale $Q$.
As we will see, this corresponds to scaling the quark masses as $m_q \rightarrow t^2 m_q$.
Similarly, $\mu_I$ must also be less than $Q$, which means that we scale it as $\mu_I\rightarrow t \mu_I$.
Following these rules, each term in the effective Lagrangian will have a well-defined chiral dimension $D$, ensuring a consistent series expansion.
The term $\Ell_{D}$ then contain all allowed terms that scale as $t^D$~\cite{weinberg_1996_vol2,WeinbergPhenom,Scherer2002IntroductionTC}.

\subsection*{Non-linear realization}

To parametrize the low energy behavior of QCD, we must find a representative element of the coset space of $G$, $G/H = \SU(2)_A$.
Let $g\in G$. 
We write $g = (U_L, U_R)$, where $U_R \in \SU(2)_R$, $U_L \in \SU(2)_L$.
Elements $h \in H$ are then of the form $h = (U, U)$.
A general element g can be written as
\begin{equation}
    g = (U_L, U_R) = (1, U_R U_L^\dagger) (U_L, U_L).
\end{equation}
Since $(U_L, U_L) \in H$, this means that we can write the coset $g H$ as $(1, U_R U_L^\dagger)H$, which gives a way to choose a representative element for each coset.
We identify
\begin{equation}
    \Sigma = U_R U_L^\dagger. 
\end{equation}
This is our standard form for elements in $gH$.
As we saw in \autoref{seciton:ccwz construction}, it therefore implicitly define transformation properties of the Goldstone bosons, which is given by the function $h(g, \xi)$.
For $\tilde g \in G$, we have 
\begin{equation}
    \tilde g (1, \Sigma)
    = (\tilde U_L, \tilde U_R) (1, U_R U_L^\dagger)
    = (1, \tilde U_R (U_R U_L^\dagger) \tilde U_L^\dagger) (\tilde U_L, \tilde U_L)
    = (1, \tilde U_R \Sigma \tilde U_L) \tilde h.
\end{equation}
This gives the transformation rule
\begin{equation}
    \Sigma \rightarrow \Sigma' = U_R \Sigma U_L^\dagger.
\end{equation}
Under transformations by $h = (U, U^\dagger) \in H$, we have
\begin{equation}
    \label{sigma transform under H}
    \Sigma \rightarrow \Sigma' = U \Sigma U^\dagger.
\end{equation}
As $\partial_\mu  (1, \Sigma) = (0, \partial_\mu \Sigma)$, the constituents of the Mauer-Cartan form are
\begin{equation}
    d_\mu = i \Sigma(x)^\dagger \partial_\mu \Sigma(x),\quad
    e_\mu = 0.
\end{equation}
Using $\partial_\mu [\Sigma(x)^\dagger\Sigma(x)] = 0 $, we can write
\begin{equation}
    d_\mu d^\mu = 
    - \Sigma(x)^\dagger [\partial_\mu \Sigma(x)] \Sigma(x)^\dagger [\partial^\mu \Sigma(x)]
    =\Sigma(x)^\dagger [\partial_\mu \Sigma(x)] [\partial^\mu \Sigma(x)^\dagger] \Sigma(x).
\end{equation}
In \autoref{seciton:ccwz construction}, we found the lowest order terms, \cref{first order terms CCWZ}.
However, as $d_\mu \in \liea{su}{2}$, which we represent by the traceless Pauli matrices, we have
\begin{equation}
    \Tr{d_\mu} = 0.
\end{equation}
This leaves us with the single leading order term
\begin{equation}
    \Tr{d_\mu d^\mu} = \Tr{\partial_\mu \Sigma (\partial^\mu \Sigma)^\dagger},
\end{equation}
where we have used the cyclic property of the trace.

However, constructing the effective Lagrangian out of terms invariant under $G$ is too restrictive to get the most general effective action.
This only allows for an even number of $d_\mu$'s, and observed processes such as the decay of the neutral pion through $\pi^0 \rightarrow \gamma \gamma$ would not be possible~\cite{Scherer2002IntroductionTC}.
This is because we have not allowed for terms that change the Lagrangian with a divergence term, as discussed in \autoref{section:symmetry}.
Terms of this type are called Wess-Zumino-Witten (WZW) terms~\cite{weinberg_1996_vol2}.
We will not consider these here, as they do not affect the thermodynamic quantities in question~\cite{Andersen:two-flavor-chpt}.

\subsection*{External currents and explicit symmetry breaking}
\label{Covarinat derivative}

Using $d_\mu$ we are able to construct any terms in the effective Lagrangian corresponding to $j=0$.
To construct the effective Lagrangian when $j \neq 0$, which is needed to capture the effects of non-zero quark masses and external currents, we treat $G =  U(1)_V \times \SU(2)_L\times \SU(2)_R$ as a gauge group, and the source currents as the corresponding gauge fields.
The effective Lagrangian is then constructed as the most general Lagrangian that is invariant under \emph{local} transformations in $G$.
The Ward-identities corresponding to the local symmetries of the Lagrangian are equivalent with the statement that $Z[j]$ is invariant under a gauge transformation of the external fields, in the absence of anomalies~\cite{Leutwyler:on_the_fundations}.

Following~\cite{Scherer2002IntroductionTC}, we write the gauge transformation as
\begin{equation}
    q_L \rightarrow e^{i\theta(x)/3} U_L(x) q_L, \quad
    q_R \rightarrow e^{i\theta(x)/3} U_R(x) q_R.
\end{equation}
First, we consider the $U(1)_V$ transformation.
The massless QCD Lagrangian then transforms as
\begin{equation}
    \Ell_\text{QCD}^0 = i \bar q \slashed{D} q
    \rightarrow
    i \bar q e^{-i\theta(x)/3} \slashed{D} e^{i\theta(x)/3} q
    = i \bar q \slashed{D} q - \frac{1}{3} \bar q \gamma^\mu q \partial_\mu \theta(x).
\end{equation}
This gives us the transformation rule
\begin{equation}
    v_{(s)}^\mu \rightarrow v_{(s)}^\mu - \partial^\mu \theta(x).
\end{equation}
Then, applying the $\SU(2)_R$ transformation, we get
\begin{equation}
    \bar q_R \slashed{D} q_R + (i \bar q_R \gamma^\mu \tau_a  q_R) \, r^a_\mu
    \rightarrow
    i \bar q_R \slashed{D} q_R + 
    (i \bar q_R \gamma^\mu q_R ) \, U_R^\dagger \partial_\mu U_R  
    + [\bar q_R \gamma^\mu (U_R^\dagger \tau_a U_r)  q_R ] \, r^a_\mu
\end{equation}
This, and the similar expression for $U_L$, gives the gauge transformation rules
\begin{align}
    r_\mu^a \tau_a & \rightarrow U_R (r_\mu^a\tau_a + i\one \partial_\mu) U_R^\dagger, \\
    l_\mu^a \tau_a & \rightarrow U_L (l_\mu^a\tau_a + i\one \partial_\mu) U_L^\dagger, \\
    s + i p & \rightarrow U_R (s + i p) U_L^\dagger, \\
    s - i p & \rightarrow U_L (s - i p) U_R^\dagger.
\end{align}
The Goldstone fields now transform as
\begin{equation}
    \Sigma(x) \rightarrow U_R(x) \Sigma(x) U_L(x)^\dagger,
\end{equation}
and the derivative as
\begin{align}
    \nonumber
    \partial_\mu \Sigma \rightarrow & \, \partial_\mu (U_R \Sigma U_L^\dagger) \\
    &= 
    U_R (\partial_\mu \Sigma )U_L^\dagger
    + (\partial_\mu  U_R) \Sigma U_L^\dagger
    + U_R \Sigma (\partial_\mu U_L^\dagger)
    \nonumber
    \\
    & = 
    U_R
    \left[
        \partial_\mu \Sigma
        + U_R^\dagger (\partial_\mu U_L) \Sigma
        + \Sigma (\partial_\mu U_L^\dagger) U_L
    \right]
    U_L^\dagger.
    \label{Sigma partial derivative}
\end{align}
The covariant derivative is defined to transform in the same way as the object it is acting upon.
This means that the definition of the covariant derivative depends on what it acts on.
Assume $A$, $B$ and $\Sigma$ transforms as $A \rightarrow U_R A U_R^\dagger$, $B \rightarrow U_L A U_L^\dagger$, and $\Sigma \rightarrow U_R \Sigma U_L^\dagger$.
In each of these cases, we define the covariant derivative as
\begin{align}
    \label{covariant derivative general}
    \nabla_\mu A &= \partial_\mu A - i [r_\mu, A], \\
    \nabla_\mu B &= \partial_\mu B - i [B, l_\mu], \\
    \nabla_\mu \Sigma &= \partial_\mu \Sigma - i r_\mu \Sigma + i \Sigma l_\mu.
\end{align}
It follows from \cref{Sigma partial derivative} that $\nabla_\mu \Sigma$ transforms as $\Sigma$, and the proof for the other two is similar.
For quantities that do not transform under $\SU(2)_L\times \SU(2)_R$, the covariant derivative is the ordinary partial derivative.
If $A$, $B$ and $C = AB$ are all quantities with a well-defined covariant derivative, then
\begin{equation}
    \nabla_\mu (AB) = (\nabla_\mu A) B + A (\nabla_\mu B).
\end{equation}
We prove the special case of $a_\mu = 0$, which is what we use in this text, in which case
\begin{equation}
    \nabla_\mu \Sigma = \partial_\mu \Sigma - i [v_\mu, \Sigma].
\end{equation}
Assume $A, B$ transforms as $\Sigma$. 
Then,
\begin{align*}
    \nabla_\mu (A B)
    & = (\partial_\mu A) B + A (\partial_\mu B) - i \com{v_\mu}{AB}
    = (\partial_\mu A - i \com{v_\mu}{A})B + A(\partial_\mu B- i \com{v_\mu}{B})\\
    & = (\nabla_\mu A)B + A (\nabla_\mu B).
\end{align*}
The more general theorem follows by applying the definition of the various covariant derivatives~\cite{Scherer:PhysRevD.53.315}.
We can extend this to do integration by parts.
Decomposing a 2-by-2 matrix $M$, as described in \autoref{section:algebra bases}, shows that the trace of the commutator of $\tau_b$ and $M$ is zero:
\begin{equation*}
    \Tr{\com{\tau_a}{M}]} = M_b\Tr{ \com{\tau_a}{\tau_b}} = 0.
\end{equation*}
Together with the fact that $\Tr{\partial_\mu A} = \partial_\mu \Tr{A}$, this gives the product rule for invariant traces:
\begin{equation*}
    \Tr{A \nabla_\mu B} = \partial_\mu \Tr{AB} - \Tr{(\nabla_\mu A) B}.
\end{equation*}
Assume $A$, $K^\mu$, and $A K^\mu$ have well-defined covariant derivatives, and that $\Tr{A K^\mu}$ is invariant under transformations by the gauge group.
Let $\Tr{K^\mu}$ be a space-time vector, and $\Tr{A}$ scalar. 
Let $\Omega$ be the domain of integration, with coordinates $x$ and $\partial \Omega$ its boundary, with coordinates $y$. Then, 
\begin{align*}
    \int_\Omega \dd x \, \Tr{A \nabla_\mu K^\mu} 
    = 
    - \int_\Omega \dd x \, \Tr{(\nabla_\mu A) K^\mu}
    + \int_{\partial\Omega} \dd y\, n_\mu \Tr{A K^\mu},
\end{align*}
where $n_\mu$ is the normal vector of $\partial \Omega$~\cite{Carroll:space-time}.
By the assumption of no variation on the boundary, the last term is constant when varying the action and may therefore be discarded.

In the presence of external source-currents, the Maurer-Cartan form becomes
\begin{equation}
    d_\mu = i \Sigma^\dagger \nabla_\mu \Sigma.
\end{equation}
Any invariant terms we constructed for $j=0$ remain invariant, however the gauge fields provide new building blocks.
The leading order kinetic term now becomes
\begin{equation}
    \label{leading order term sigma}
    \Tr{d_\mu d^\mu} = \Tr{\nabla_\mu \Sigma (\nabla^\mu \Sigma)^\dagger}
\end{equation}

We define the scalar current
\begin{equation}
    \chi = 2 B_0 (s + ip),
\end{equation}
where $B_0$ is defined by the up quark condensate in the chiral limit by
\begin{equation}
    \ex{\bar u u}_{j = 0} = - f^2 B_0.
\end{equation}
Here, $f$ is the bare pion decay constant.
The pseudoscalar current is thus $\chi^\dagger = 2 B_0 (s - ip)$.
As we are going to set $\chi = 2 B_0 m$, both $\chi$ and $\chi^\dagger$ are of chiral dimension $2$.
These can be combined to give more gauge-invariant terms, such as
\begin{equation}
    \Tr{\chi^\dagger \Sigma}, \quad \Tr{\Sigma^\dagger \chi}.
\end{equation}
However, under parity transformation, $\chi \rightarrow \chi^\dagger$ and $\Sigma \rightarrow \Sigma^\dagger$~\cite{Scherer2002IntroductionTC}, so to ensure the Lagrangian is a scalar, the only allowed term is\footnote{This corresponds to the fact that pions are pseudo scalars, they transform as $\pi \rightarrow - \pi$ under parity, as they are exitations in $\SU(2)_A$.}
\begin{equation}
    \label{leading order chi sigma term}
    \Tr{\chi^\dagger \Sigma + \Sigma^\dagger \chi}
\end{equation}

In the grand canonical ensemble, as described in \cref{section:statistical mechanics}, we modify the Lagrangian as
\begin{equation}
    \Ell \rightarrow \Ell + \mu Q,
\end{equation}
where $Q$ is a conserved charge, and $\mu$ is the corresponding chemical component.
We are interested systems with non-zero isospin, 
\begin{equation}
    Q_I = \int \dd^3 x \, V^0_3 = \int \dd^3 x \, \frac{1}{2}  \bar q \gamma^0 \tau_3 q,
\end{equation}
and the corresponding isospin chemical potential $\mu_I$.
We do this by considering the system with a external current
\begin{equation}
    v_\mu = \frac{1}{2} \mu_I \delta_\mu^0 \tau_3.
\end{equation}
The covariant derivative therefore is of chiral dimension 1.
As in the case of Yang-Mills theory, we may form a field strength tensor to create invariant terms, in this case they are
\begin{align}
    f_{\mu\nu}^R &= \partial_\mu r_\nu - \partial_\nu r_\mu - i [r_\mu, r_\nu], \\
    f_{\mu\nu}^L &= \partial_\mu l_\nu - \partial_\nu l_\mu - i [l_\mu, l_\nu].
\end{align}
In our case, these vanish, and we can therefore safely ignore terms made including them.

For $j = 0$, the ground state is by assumption $\Sigma = \one$, the vacuum, and we can use the paramterization
\begin{equation}
    \label{vacuum parametrization}
    \Sigma(x) = \exp{i \frac{\pi_a\tau_a}{f}},
\end{equation}
where $f$ is the bare pion decay constant, $\pi_a$ are the three Goldstone bosons, a set of real functions of space-time.
This ensures that $\pi = 0$ corresponds to the vacuum.
If we perform an infinitesimal isospin-transformation, and assume $\pi_a/f$ small, then
\begin{equation}
    \Sigma \rightarrow U_V \Sigma U_V^\dagger
    =
    \left(1 + i \eta_a \frac{1}{2} \tau_a\right)
    \left(1 + i \frac{1}{f} \pi_b  \tau_b\right)
    \left(1 - i \eta_c \frac{1}{2} \tau_c\right)
    =
    1 + i\frac{1}{f}\pi_a (\delta_{ac} + i \eta_b \epsilon_{abc}) \tau_c,
\end{equation}
or
\begin{equation}
    \pi_a \rightarrow (\delta_{ac} + i \eta_b \epsilon_{abc}) \pi_c.
\end{equation}
The generators of $\pi_a$ under isospin-transforamtions are thus the adjoint representation of $\mathfrak{su}(2)$, and they form an isospin triplet.
For $\eta_1 = \eta_2 = 0$, i.e. transformations generated by $\tau_3$, $\pi_3$ is invariant, which means that it has quantum number $I_3 = 0$.\footnote{Authors differe if they define $\sqrt 2 \pi_\pm = \pi_1 \pm i \pi_2$, or with opposite signs.We choose the former, so that $\pi_+ \ket{0}$ is the state with the quantum numbers of the positive pion.}
$\pi_1$ and $\pi_2$ do not have a definite value of the third component of isospin, but rather for the first and second component.
They are related to the observed, charged pions $\pi_+$ and $\pi_-$ by~\cite{Scherer2002IntroductionTC}
\begin{equation}
    \pi_a\tau_a
    = 
    \begin{pmatrix}
        \pi_3 & \pi_1 - i \pi_2 \\
        \pi_1 + i \pi_2 & - \pi_3
    \end{pmatrix}
    = 
    \begin{pmatrix}
        \pi_0 & \sqrt{2} \pi_- \\
        \sqrt 2 \pi_+ & - \pi_0
    \end{pmatrix},
\end{equation}
where $\pi_\pm$ has a third isospin-component of $I_3 = \pm1$.
For non-zero isospin chemical potential, however, we expect that the ground state may be rotated away from the vacuum.
To find what the new groundstate is, we have to minimize the Hamiltonian.

