\section{CCSW construction}

As Goldstone bosons are massless, they play a crucial role in the low energy dynamics.
To best describe this limit, we seek a parametrization of the fields in which they are the degrees of freedom.
We saw that the Goldstone bosons corresponds to excitations within the vacuum manifold.
The vacuum manifold corresponds to points in field space $\varphi$ that can be reach from the vacuum $\varphi_0$ from a transformation $g \in G$.
This means that we can write such excitations as
\begin{equation}
    \varphi(x) = \tilde\Sigma(x) \varphi_0, \quad \tilde \Sigma(x) = \exp{i \eta_\alpha T_\alpha}
\end{equation}
The real functions $\eta_\alpha(x)$ thus parametrizes fluctuations away from the ground state $\varphi_0$ in to the ground state manifold.
%TODO: Forklar hvorfor vi ikke bryr oss
% $\varphi(x)$ is not a completely generic field configurations, as it does not account for fluctuations in directions other than those in $G$, but we don't care. (?)
This parametrization, however, is highly redundant.
% We may write any $\Sigma$ as
% \begin{equation}
%     \Sigma = \exp{i \xi_a t_a} = \exp{i \tilde \xi_a(\xi) \tilde t_a} \exp{i \hat \xi_a(\xi) \hat t_a}
% \end{equation}
Two elements $\tilde\Sigma$ and $\tilde\Sigma'$, related by 
\begin{equation}
    \tilde \Sigma' = \tilde\Sigma e^{i \theta_a t_a}
\end{equation}
results in the same $\varphi(x)$.
This is because  $e^{i \theta_a t_a} = h \in H$, and $h \varphi_0 = \varphi_0$, by assumption.
The set of all equivalent $\Sigma$'s is exactly the left coset, $gH = \setbuilder{gh}{ h \in H}$.
The set of cosets forms a new manifold, $G / H$, called the Goldstone manifold, and has dimension $\dim(G/H) = \dim(G) - \dim(H)$.
The cosets $gH$ form a partition of $G$, as it is an equivalence relation. ($g \sim g'$ if $g' = gh$.)
This means that each element of $G$ belongs to one, and only one, coset.
To remove the redundancy in the parametrization, we need to choose one representative element from each coset.
A general element, $g \in G$ can be written uniquely as
\begin{equation}
    g = \exp{i \xi_i x_i} \exp{i \theta_a t_a}.
\end{equation}
(HVORFOR? FORKLAR!)
We therefore define a map
\begin{align}
    f: G/H & \longrightarrow G, \\
    gH & \longrightarrow \Sigma_g, \\
    & g = \exp{i \xi_i x_i} \exp{i \theta_a t_a}, \quad \Sigma_g = \exp{i \xi_i x_i}.
\end{align}

This means that the generic point in field space, $\varphi$, can be written
\begin{equation*}
    \varphi = \Sigma \varphi_0, \quad \Sigma \in G/H.
\end{equation*}
% The coset is itself a Lie group, and thus also a set of generators $\{\tau_a\}$, which span the corresponding Lie algebra. %! NO!!!


In the linear sigma model, the original $O(N)$ symmetry is broken down to $O(N-1)$, which transforms the remaining $N-1$ fields with vanishing expectation value into each other.
However, $O(N)$ consists of two disconnected subsets, those matrices with determinant 1 and those with determinant -1.
There is no continuous path that takes an element of $O(N)$ with determinant of $-1$.\footnote{A simple proof of this is the fact that the determinant is a continuous function, while any path $\det A(t)$ such that $\det A(1) = -1,\, \det A(0) = 1$ must make a discontinuous jump.}
The set of symmetries that are connected to the identity is
\begin{eqnarray}
    G = SO(N) = \setbuilder{M \in O(N)}{ \det M = 1},
\end{eqnarray}
which is broken down to $SO(N-1)$.
The Goldstone manifold is thus $G/H = SO(N) / SO(N-1)$.

\subsection*{Transformation properties of Goldstone bosons}
We can deduce how the Goldstone bosons transforms under $G$ from how $\varphi$ transforms.
In general, 
\begin{equation}
    \varphi'(x) = g \varphi(x) = (g \Sigma(x)) \varphi_0 = \Sigma'(x) \varphi_0 \quad g \in G.
\end{equation}
Now, in general $g \Sigma(x)$ is not in the standard form $\Sigma'(x) = \exp{i \pi'_a(x)\tau^a}$, as they can differ by an element of the remaining symmetry group, $g\Sigma(x) = \Sigma'(x) h[g](x)$, $h[g](x) \in H$.
Therefore, we get the transformation rule
\begin{equation}
    \Sigma' = g \Sigma (h[g])^{-1}.
\end{equation}
